---
title: "Figures for: Communicating Uncertainty Across Disciplines"
author: "Bahman Rostami-Tabar"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    theme: cosmo
  pdf:
    toc: true
    keep-tex: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
# Install missing packages if needed:
# install.packages(c("tidyverse", "gt", "scales", "patchwork", "ggtext"))
library(tidyverse)
library(gt)
library(scales)
library(patchwork)

# Common theme for all ggplot figures
theme_uncertainty <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid.minor   = element_blank(),
      plot.title         = element_text(face = "bold", size = base_size + 1),
      plot.subtitle      = element_text(color = "grey40", size = base_size - 1),
      plot.caption       = element_text(color = "grey50", size = base_size - 2,
                                        hjust = 0),
      axis.title         = element_text(face = "bold"),
      legend.position    = "bottom"
    )
}
```

# Section 2 Figures {#sec-fig-2}

## Figure 2.1 — Stirling's Four Dimensions of Incertitude {#fig-stirling}

```{r fig-stirling}
#| label: fig-stirling
#| fig-cap: "Stirling's four dimensions of incertitude, structured by whether
#|   knowledge about outcomes and knowledge about likelihoods are problematic
#|   or unproblematic. These are states of knowledge, not states of nature.
#|   Adapted from Stirling (n.d.) and Scoones (2019)."
#| fig-width: 7
#| fig-height: 6

quadrants <- tibble(
  xmin  = c(0,   0.5, 0,   0.5),
  xmax  = c(0.5, 1,   0.5, 1),
  ymin  = c(0,   0,   0.5, 0.5),
  ymax  = c(0.5, 0.5, 1,   1),
  fill  = c("#2ecc71", "#e67e22", "#3498db", "#c0392b"),
  label = c("RISK", "AMBIGUITY", "UNCERTAINTY", "IGNORANCE"),
  desc  = c(
    "Outcomes known\nLikelihoods estimable",
    "Outcomes contested\nLikelihoods estimable",
    "Outcomes known\nLikelihoods unknown",
    "Outcomes unknown\nLikelihoods unknown"
  ),
  lx = c(0.25, 0.75, 0.25, 0.75),
  ly = c(0.25, 0.25, 0.75, 0.75)
)

ggplot(quadrants) +
  geom_rect(
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill),
    alpha = 0.22, colour = "grey50", linewidth = 0.6
  ) +
  scale_fill_identity() +
  geom_text(
    aes(x = lx, y = ly + 0.09, label = label),
    fontface = "bold", size = 5.5
  ) +
  geom_text(
    aes(x = lx, y = ly - 0.07, label = desc),
    size = 3.3, colour = "grey30", lineheight = 1.3
  ) +
  # Axis tick labels
  annotate("text", x = 0.25, y = -0.06, label = "Unproblematic",
           size = 3.4, colour = "grey45", fontface = "italic") +
  annotate("text", x = 0.75, y = -0.06, label = "Problematic",
           size = 3.4, colour = "grey45", fontface = "italic") +
  annotate("text", x = -0.07, y = 0.25, label = "Unproblematic",
           size = 3.4, colour = "grey45", fontface = "italic", angle = 90) +
  annotate("text", x = -0.07, y = 0.75, label = "Problematic",
           size = 3.4, colour = "grey45", fontface = "italic", angle = 90) +
  # Axis titles
  annotate("text", x = 0.5, y = -0.13, label = "Knowledge about Outcomes",
           size = 4.2, fontface = "bold") +
  annotate("text", x = -0.15, y = 0.5, label = "Knowledge about Likelihoods",
           size = 4.2, fontface = "bold", angle = 90) +
  coord_fixed(xlim = c(-0.2, 1.05), ylim = c(-0.2, 1.05), clip = "off") +
  theme_void() +
  theme(plot.margin = margin(10, 10, 40, 50)) +
  labs(
    caption = "Source: Adapted from Stirling (n.d.) and Scoones (2019)"
  )
```

## Figure 2.2 — Known and Unknown Unknowns {#fig-knowns}

```{r fig-knowns}
#| label: fig-knowns
#| fig-cap: "The four-quadrant taxonomy of known and unknown unknowns, adapted
#|   from Makridakis and Bakas (2016). Each quadrant implies a different
#|   forecasting approach and level of associated risk."
#| fig-width: 8
#| fig-height: 6.5

knowns <- tibble(
  xmin  = c(0,   0.5, 0,   0.5),
  xmax  = c(0.5, 1,   0.5, 1),
  ymin  = c(0,   0,   0.5, 0.5),
  ymax  = c(0.5, 0.5, 1,   1),
  fill  = c("#27ae60", "#f39c12", "#2980b9", "#c0392b"),
  label = c("Known / Knowns", "Unknown / Knowns",
            "Known / Unknowns", "Unknown / Unknowns"),
  desc  = c(
    "Normal conditions\nRisk estimable\nThin-tailed errors",
    "Knowns that are\nignored or denied\nRisk underestimated",
    "Awareness of ignorance\nFat-tailed errors\nRisk: hard to estimate",
    "Black Swans\nEntirely unpredictable\nRisk: infinite"
  ),
  lx = c(0.25, 0.75, 0.25, 0.75),
  ly = c(0.25, 0.25, 0.75, 0.75)
)

ggplot(knowns) +
  geom_rect(
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill),
    alpha = 0.22, colour = "grey50", linewidth = 0.6
  ) +
  scale_fill_identity() +
  geom_text(
    aes(x = lx, y = ly + 0.10, label = label),
    fontface = "bold", size = 4.5
  ) +
  geom_text(
    aes(x = lx, y = ly - 0.06, label = desc),
    size = 3.1, colour = "grey30", lineheight = 1.3
  ) +
  # Axis tick labels
  annotate("text", x = 0.25, y = -0.06, label = "Known",
           size = 3.5, colour = "grey40", fontface = "italic") +
  annotate("text", x = 0.75, y = -0.06, label = "Unknown",
           size = 3.5, colour = "grey40", fontface = "italic") +
  annotate("text", x = -0.07, y = 0.25, label = "Known",
           size = 3.5, colour = "grey40", fontface = "italic", angle = 90) +
  annotate("text", x = -0.07, y = 0.75, label = "Unknown",
           size = 3.5, colour = "grey40", fontface = "italic", angle = 90) +
  # Axis titles
  annotate("text", x = 0.5, y = -0.13, label = "Outcomes",
           size = 4.5, fontface = "bold") +
  annotate("text", x = -0.15, y = 0.5, label = "Likelihoods / Mechanisms",
           size = 4.5, fontface = "bold", angle = 90) +
  coord_fixed(xlim = c(-0.22, 1.05), ylim = c(-0.22, 1.05), clip = "off") +
  theme_void() +
  theme(plot.margin = margin(10, 10, 40, 55)) +
  labs(
    caption = "Source: Adapted from Makridakis and Bakas (2016)"
  )
```

# Section 4 Figures {#sec-fig-4}

## Figure 4.1 — Concept Comparison Table {#tbl-concepts}

```{r tbl-concepts}
#| label: tbl-concepts
#| tbl-cap: "Key distinctions between uncertainty and four related concepts,
#|   across dimensions relevant to uncertainty communication practice."

tibble(
  Concept       = c("Uncertainty", "Risk", "Variability",
                    "Randomness", "Volatility"),
  Definition    = c(
    "Unknown probabilities of outcomes",
    "Known probabilities of known outcomes",
    "Real differences within a population or system",
    "Irreducible stochastic variation (aleatory)",
    "Measurable fluctuation in financial/economic time series"
  ),
  Knowability   = c(
    "Probabilities unknown",
    "Fully quantifiable",
    "Directly measurable",
    "Statistically characterisable",
    "Estimable (e.g., GARCH models)"
  ),
  Reducibility  = c(
    "Partially, through better models or data",
    "Already quantified",
    "Not reducible — it is real",
    "Irreducible — structural property",
    "Model-dependent"
  ),
  `Primary\ncommunication challenge` = c(
    "Avoiding false precision; resisting collapse into risk",
    "Ensuring calibration",
    "Distinguishing from epistemic uncertainty",
    "Avoiding deterministic misinterpretation",
    "Distinguishing from deep/structural uncertainty"
  )
) |>
  gt() |>
  tab_header(
    title    = "Uncertainty and Related Concepts: Key Distinctions",
    subtitle = "Synthesised from Scoones (2019), Stirling (n.d.), Eggen (2025), and EFSA (2019)"
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style     = cell_text(weight = "bold", color = "#1a5276"),
    locations = cells_body(columns = Concept)
  ) |>
  tab_style(
    style     = cell_fill(color = "#eaf4fb"),
    locations = cells_body(rows = c(1, 3, 5))
  ) |>
  tab_style(
    style     = cell_fill(color = "#fdfefe"),
    locations = cells_body(rows = c(2, 4))
  ) |>
  cols_width(
    Concept       ~ px(110),
    Definition    ~ px(210),
    Knowability   ~ px(145),
    Reducibility  ~ px(175),
    `Primary\ncommunication challenge` ~ px(200)
  ) |>
  tab_options(
    table.font.size          = 12,
    heading.title.font.size  = 13,
    heading.subtitle.font.size = 11,
    column_labels.background.color = "#2c3e50",
    column_labels.font.color       = "white"
  )
```

# Section 5 — Method Selection Flowchart {#sec-fig-5}

## Figure 5.1 — How to Choose a Communication Method {#fig-flowchart}

```{mermaid}
%%| label: fig-flowchart
%%| fig-cap: "Decision flowchart for selecting an uncertainty communication
%%|   method, based on audience level, type of uncertainty, and decision
%%|   context. Synthesised from the review in Section 6."

flowchart TD
    A([Start]) --> B{Is the uncertainty\nquantifiable with\nreliable probabilities?}

    B -->|No| C{What is the\nprimary goal?}
    B -->|Yes| D{What is the\naudience level?}

    C -->|Express evidence\nquality / confidence| E[🟡 Categorical Approach\nor Verbal Expression]
    C -->|Explore plausible\nfutures| F[🟠 Scenario /\nWhat-If Analysis]

    D -->|Entry-level\nGeneral public| G[🟢 Numerical Expression\nor Simple Interval]
    D -->|Informed\nPolicymakers / journalists| H{Does the decision\nneed a specific\nthreshold?}
    D -->|Technical\nExperts / analysts| I{Does the decision\ndepend on cumulative\nor path outcomes?}

    H -->|Yes| J[🟢 Quantile Forecast\nor Prediction Interval]
    H -->|No| K[🟡 Verbal + Numeric\nor Categorical Scale]

    I -->|Yes| L[🔵 Forecast Trajectories\nEnsemble / Spaghetti Plot]
    I -->|No| M{How much\ndistributional\ndetail is needed?}

    M -->|Full shape| N[🔵 Probability Distribution\nPDF / CDF / CCDF]
    M -->|Time-varying\nwith horizon| O[🔵 Fan Chart]
    M -->|Key quantiles\nonly| P[🟢 Prediction Intervals\nand Quantiles]

    style A fill:#2c3e50,color:#fff,stroke:none
    style E fill:#f39c12,color:#fff,stroke:none
    style F fill:#e67e22,color:#fff,stroke:none
    style G fill:#27ae60,color:#fff,stroke:none
    style J fill:#27ae60,color:#fff,stroke:none
    style K fill:#f39c12,color:#fff,stroke:none
    style L fill:#2980b9,color:#fff,stroke:none
    style N fill:#2980b9,color:#fff,stroke:none
    style O fill:#2980b9,color:#fff,stroke:none
    style P fill:#27ae60,color:#fff,stroke:none
```

# Section 6 Figures — Communication Method Illustrations {#sec-fig-6}

## Figure 6.1 — Forecast Trajectories {#fig-trajectories}

```{r fig-trajectories}
#| label: fig-trajectories
#| fig-cap: "Illustrative forecast trajectory (spaghetti) plot. Each coloured
#|   line represents one possible future path from an ensemble model. The dashed
#|   line shows the central (median) forecast. The vertical dotted line marks
#|   the forecast origin."
#| fig-width: 8
#| fig-height: 5

set.seed(123)
n_obs  <- 24
n_fore <- 20
n_traj <- 10

obs_data <- tibble(
  time  = 1:n_obs,
  value = 10 + cumsum(rnorm(n_obs, 0.25, 0.7))
)

last_val <- tail(obs_data$value, 1)

traj_data <- map_dfr(seq_len(n_traj), function(i) {
  tibble(
    time       = n_obs:(n_obs + n_fore),
    value      = c(last_val,
                   last_val + cumsum(rnorm(n_fore, 0.2, 0.9))),
    trajectory = factor(i)
  )
})

central_fore <- tibble(
  time  = n_obs:(n_obs + n_fore),
  value = last_val + c(0, cumsum(rep(0.2, n_fore)))
)

ggplot() +
  geom_line(
    data = traj_data,
    aes(x = time, y = value, group = trajectory),
    colour = "#2980b9", alpha = 0.45, linewidth = 0.7
  ) +
  geom_line(
    data = obs_data,
    aes(x = time, y = value),
    colour = "black", linewidth = 1.3
  ) +
  geom_line(
    data = central_fore,
    aes(x = time, y = value),
    colour = "#c0392b", linewidth = 1.2, linetype = "dashed"
  ) +
  geom_vline(
    xintercept = n_obs, linetype = "dotted", colour = "grey50", linewidth = 0.8
  ) +
  annotate("text", x = n_obs / 2, y = max(obs_data$value) + 0.6,
           label = "Observed", size = 3.5, colour = "grey30") +
  annotate("text", x = n_obs + n_fore / 2,
           y = max(traj_data$value) + 0.7,
           label = "Ensemble trajectories", size = 3.5, colour = "#2980b9") +
  annotate("text", x = n_obs + 1,
           y = tail(central_fore$value, 1) + 0.5,
           label = "Median forecast", size = 3.2, colour = "#c0392b",
           hjust = 0) +
  scale_x_continuous(breaks = seq(0, n_obs + n_fore, by = 5)) +
  theme_uncertainty() +
  labs(
    title    = "Forecast Trajectories",
    subtitle = paste0("Multiple ensemble members (n = ", n_traj,
                      ") showing the range of possible futures"),
    x = "Time", y = "Value",
    caption  = "Illustrative example. Each line represents one simulated future path."
  )
```

## Figure 6.2 — Fan Chart {#fig-fan}

```{r fig-fan}
#| label: fig-fan
#| fig-cap: "Illustrative fan chart showing a central forecast surrounded by
#|   nested probability bands at 30%, 50%, 70%, and 90% confidence levels.
#|   Darker shading indicates higher probability density near the central
#|   forecast."
#| fig-width: 8
#| fig-height: 5

set.seed(42)
n_obs  <- 20
n_fore <- 30

obs_data <- tibble(
  time  = 1:n_obs,
  value = 10 + cumsum(rnorm(n_obs, 0.2, 0.6))
)

last_val <- tail(obs_data$value, 1)
h        <- seq_len(n_fore)
central  <- last_val + cumsum(rep(0.2, n_fore))
spread   <- sqrt(h) * 0.55  # fan widens with horizon

fore_data <- tibble(
  time    = n_obs + h,
  central = central,
  lo90 = central - qnorm(0.95) * spread,
  hi90 = central + qnorm(0.95) * spread,
  lo70 = central - qnorm(0.85) * spread,
  hi70 = central + qnorm(0.85) * spread,
  lo50 = central - qnorm(0.75) * spread,
  hi50 = central + qnorm(0.75) * spread,
  lo30 = central - qnorm(0.65) * spread,
  hi30 = central + qnorm(0.65) * spread
)

# Legend data
band_legend <- tibble(
  label = c("90%", "70%", "50%", "30%"),
  alpha = c(0.12, 0.18, 0.25, 0.35),
  fill  = rep("#2980b9", 4)
)

ggplot() +
  geom_ribbon(data = fore_data, aes(x = time, ymin = lo90, ymax = hi90),
              fill = "#2980b9", alpha = 0.12) +
  geom_ribbon(data = fore_data, aes(x = time, ymin = lo70, ymax = hi70),
              fill = "#2980b9", alpha = 0.18) +
  geom_ribbon(data = fore_data, aes(x = time, ymin = lo50, ymax = hi50),
              fill = "#2980b9", alpha = 0.25) +
  geom_ribbon(data = fore_data, aes(x = time, ymin = lo30, ymax = hi30),
              fill = "#2980b9", alpha = 0.35) +
  geom_line(data = obs_data, aes(x = time, y = value),
            colour = "black", linewidth = 1.3) +
  geom_line(data = fore_data, aes(x = time, y = central),
            colour = "#c0392b", linewidth = 1.1, linetype = "dashed") +
  geom_vline(xintercept = n_obs, linetype = "dotted",
             colour = "grey50", linewidth = 0.8) +
  # Band labels at the end of the forecast
  annotate("text",
           x     = n_obs + n_fore + 0.5,
           y     = c(tail(fore_data$hi90, 1), tail(fore_data$hi70, 1),
                     tail(fore_data$hi50, 1), tail(fore_data$hi30, 1)),
           label = c("90%", "70%", "50%", "30%"),
           size  = 3.2, colour = "#1a5276", hjust = 0) +
  scale_x_continuous(
    limits = c(1, n_obs + n_fore + 4),
    breaks = seq(0, n_obs + n_fore, by = 5)
  ) +
  theme_uncertainty() +
  labs(
    title    = "Fan Chart",
    subtitle = "Central forecast (dashed) with probability bands at 30%, 50%, 70%, and 90%",
    x = "Time", y = "Value",
    caption  = "Illustrative example. Bands widen with the forecast horizon."
  )
```

## Figure 6.3 — Probability Distribution {#fig-distribution}

```{r fig-distribution}
#| label: fig-distribution
#| fig-cap: "Illustrative probability density function (PDF) showing the full
#|   distribution of possible outcomes. The red shaded region highlights the
#|   upper tail above a critical threshold — the region of primary concern for
#|   risk management."
#| fig-width: 8
#| fig-height: 5

x <- seq(-4, 5, length.out = 600)

dist_data <- tibble(
  x       = x,
  density = dnorm(x, mean = 0.5, sd = 1.2)
)

threshold <- 2.5
mean_val  <- 0.5

ggplot(dist_data, aes(x = x, y = density)) +
  geom_ribbon(
    data = filter(dist_data, x < threshold),
    aes(ymin = 0, ymax = density),
    fill = "#2980b9", alpha = 0.18
  ) +
  geom_ribbon(
    data = filter(dist_data, x >= threshold),
    aes(ymin = 0, ymax = density),
    fill = "#c0392b", alpha = 0.45
  ) +
  geom_line(linewidth = 1.3, colour = "#1a5276") +
  geom_vline(xintercept = mean_val, linetype = "dashed",
             colour = "grey40", linewidth = 0.9) +
  geom_vline(xintercept = threshold, linetype = "dashed",
             colour = "#c0392b", linewidth = 0.9) +
  # Annotations
  annotate("text", x = mean_val, y = max(dist_data$density) + 0.015,
           label = "Mean", size = 3.5, colour = "grey40", hjust = 0.5) +
  annotate("text", x = threshold + 0.1,
           y = max(dist_data$density) + 0.015,
           label = "Threshold", size = 3.5, colour = "#c0392b", hjust = 0) +
  annotate("text", x = 3.5, y = 0.03,
           label = "High-risk\nregion", size = 3.2, colour = "#c0392b",
           lineheight = 1.2) +
  annotate("text", x = -2, y = 0.06,
           label = "Most probable\noutcomes", size = 3.2, colour = "#2980b9",
           lineheight = 1.2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08))) +
  theme_uncertainty() +
  labs(
    title    = "Probability Distribution (PDF)",
    subtitle = "Full distribution of outcomes with critical threshold and risk region marked",
    x = "Outcome", y = "Probability Density",
    caption  = "Illustrative example. Area under the curve sums to 1."
  )
```

## Figure 6.4 — Prediction Intervals {#fig-intervals}

```{r fig-intervals}
#| label: fig-intervals
#| fig-cap: "Illustrative prediction intervals showing an observed time series
#|   followed by a central forecast with 80% and 95% prediction intervals.
#|   Wider intervals reflect greater uncertainty at longer forecast horizons."
#| fig-width: 8
#| fig-height: 5

set.seed(99)
n_obs  <- 20
n_fore <- 25

obs_data <- tibble(
  time  = 1:n_obs,
  value = 10 + cumsum(rnorm(n_obs, 0.25, 0.6))
)

last_val <- tail(obs_data$value, 1)
h        <- seq_len(n_fore)
central  <- last_val + cumsum(rep(0.22, n_fore))
spread   <- sqrt(h) * 0.5

fore_data <- tibble(
  time    = n_obs + h,
  central = central,
  lo95    = central - qnorm(0.975) * spread,
  hi95    = central + qnorm(0.975) * spread,
  lo80    = central - qnorm(0.90)  * spread,
  hi80    = central + qnorm(0.90)  * spread
)

ggplot() +
  geom_ribbon(data = fore_data, aes(x = time, ymin = lo95, ymax = hi95),
              fill = "#2980b9", alpha = 0.18) +
  geom_ribbon(data = fore_data, aes(x = time, ymin = lo80, ymax = hi80),
              fill = "#2980b9", alpha = 0.35) +
  geom_line(data = obs_data, aes(x = time, y = value),
            colour = "black", linewidth = 1.3) +
  geom_line(data = fore_data, aes(x = time, y = central),
            colour = "#c0392b", linewidth = 1.1, linetype = "dashed") +
  geom_vline(xintercept = n_obs, linetype = "dotted",
             colour = "grey50", linewidth = 0.8) +
  # Interval labels
  annotate("text",
           x     = n_obs + n_fore - 1,
           y     = c(tail(fore_data$hi95, 1) + 0.25,
                     tail(fore_data$hi80, 1) + 0.25),
           label = c("95% PI", "80% PI"),
           size  = 3.2, colour = "#1a5276", hjust = 0.5) +
  annotate("text",
           x = n_obs / 2, y = max(obs_data$value) + 0.5,
           label = "Observed", size = 3.5, colour = "grey30") +
  scale_x_continuous(breaks = seq(0, n_obs + n_fore, by = 5)) +
  theme_uncertainty() +
  labs(
    title    = "Prediction Intervals",
    subtitle = "Central forecast (dashed) with 80% and 95% prediction intervals (PI)",
    x = "Time", y = "Value",
    caption  = paste0("Illustrative example. Intervals widen with forecast ",
                      "horizon, reflecting accumulating uncertainty.")
  )
```

## Figure 6.5 — Categorical (RAG) Approach {#fig-rag}

```{r fig-rag}
#| label: fig-rag
#| fig-cap: "Illustrative Red-Amber-Green (RAG) categorical uncertainty
#|   display for five indicators. Each indicator is assigned a category
#|   (Low, Moderate, High) based on the underlying uncertainty assessment."
#| fig-width: 7
#| fig-height: 4

rag_data <- tibble(
  Indicator   = c("Indicator A", "Indicator B", "Indicator C",
                  "Indicator D", "Indicator E"),
  Category    = c("Low", "Low", "Moderate", "High", "High"),
  Probability = c("< 10%", "5–15%", "25–45%", "55–70%", "> 75%"),
  Confidence  = c("High", "High", "Moderate", "Low", "Low")
) |>
  mutate(
    Category  = factor(Category, levels = c("Low", "Moderate", "High")),
    Indicator = fct_rev(factor(Indicator))
  )

rag_colors <- c("Low" = "#27ae60", "Moderate" = "#f39c12", "High" = "#c0392b")

ggplot(rag_data, aes(x = 1, y = Indicator, fill = Category)) +
  geom_tile(colour = "white", linewidth = 2, width = 0.6, height = 0.7) +
  geom_text(aes(label = paste0(Category, "\n(", Probability, ")")),
            size = 3.2, colour = "white", fontface = "bold") +
  scale_fill_manual(values = rag_colors) +
  theme_uncertainty() +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    panel.grid   = element_blank(),
    legend.position = "right"
  ) +
  labs(
    title    = "Categorical (RAG) Uncertainty Display",
    subtitle = "Five indicators classified by uncertainty level with indicative probability ranges",
    y        = NULL,
    fill     = "Uncertainty\nLevel",
    caption  = "Illustrative example. Cut-offs should be explicitly defined in practice."
  )
```

## Figure 6.6 — Scenario Analysis {#fig-scenarios}

```{r fig-scenarios}
#| label: fig-scenarios
#| fig-cap: "Illustrative scenario analysis showing four distinct plausible
#|   futures without explicit probabilities. An even number of scenarios
#|   prevents audiences from defaulting to a middle option as a deterministic
#|   prediction."
#| fig-width: 8
#| fig-height: 5

set.seed(55)
n_obs  <- 15
n_fore <- 25

obs_data <- tibble(
  time  = 1:n_obs,
  value = 10 + cumsum(rnorm(n_obs, 0.2, 0.5))
)

last_val <- tail(obs_data$value, 1)
t_fore   <- (n_obs + 1):(n_obs + n_fore)

scenarios <- tibble(
  time = rep(t_fore, 4),
  value = c(
    last_val + cumsum(rnorm(n_fore, 0.55, 0.25)),   # High growth
    last_val + cumsum(rnorm(n_fore, 0.25, 0.20)),   # Moderate growth
    last_val + cumsum(rnorm(n_fore, -0.05, 0.25)),  # Stagnation
    last_val + cumsum(rnorm(n_fore, -0.40, 0.30))   # Decline
  ),
  scenario = rep(c("S1: High growth", "S2: Moderate growth",
                   "S3: Stagnation", "S4: Decline"), each = n_fore)
)

scenario_colors <- c(
  "S1: High growth"     = "#27ae60",
  "S2: Moderate growth" = "#2980b9",
  "S3: Stagnation"      = "#f39c12",
  "S4: Decline"         = "#c0392b"
)

ggplot() +
  geom_line(data = obs_data, aes(x = time, y = value),
            colour = "black", linewidth = 1.3) +
  geom_line(
    data = scenarios,
    aes(x = time, y = value, colour = scenario, linetype = scenario),
    linewidth = 1.1
  ) +
  geom_vline(xintercept = n_obs, linetype = "dotted",
             colour = "grey50", linewidth = 0.8) +
  scale_colour_manual(values = scenario_colors) +
  scale_linetype_manual(
    values = c("solid", "dashed", "dotdash", "longdash")
  ) +
  theme_uncertainty() +
  theme(legend.title = element_blank()) +
  labs(
    title    = "Scenario Analysis",
    subtitle = "Four distinct plausible futures — no probabilities assigned to individual scenarios",
    x = "Time", y = "Value",
    caption  = paste0(
      "Illustrative example. Four scenarios shown to prevent anchoring on a ",
      "'middle' option.\n",
      "Good practice: Stirling (n.d.) recommends an even number of scenarios."
    )
  )
```

## Figure 6.7 — Verbal Probability Scale {#tbl-verbal}

```{r tbl-verbal}
#| label: tbl-verbal
#| tbl-cap: "Illustrative verbal probability scale mapping qualitative
#|   expressions to approximate numerical probability ranges, based on
#|   frameworks used by the IPCC and intelligence agencies. The rightmost
#|   column shows the average spread in public interpretation of each term."

tibble(
  `Verbal expression`   = c("Virtually certain", "Very likely", "Likely",
                             "About as likely as not", "Unlikely",
                             "Very unlikely", "Exceptionally unlikely"),
  `Approximate range`   = c("99–100%", "90–100%", "66–100%",
                             "33–66%",  "0–33%",   "0–10%", "0–1%"),
  `IPCC term`           = c("✓", "✓", "✓", "✓", "✓", "✓", "✓"),
  `Typical public\ninterpretation spread` = c(
    "Narrow", "Narrow", "Moderate",
    "Wide", "Wide", "Wide", "Very wide (0% anchor risk)"
  )
) |>
  gt() |>
  tab_header(
    title    = "Verbal Probability Scale with Numeric Translations",
    subtitle = "Based on IPCC (Mastrandrea et al., 2010) and Dhami & Mandel (2022)"
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style     = cell_fill(color = "#eaf4fb"),
    locations = cells_body(rows = c(1, 3, 5, 7))
  ) |>
  tab_style(
    style     = cell_text(color = "#c0392b"),
    locations = cells_body(
      columns = `Typical public\ninterpretation spread`,
      rows    = c(5, 6, 7)
    )
  ) |>
  tab_source_note(
    "Note: Always pair verbal expressions with numeric translations in practice.
     Present the number before the word (e.g., '10% — unlikely')."
  ) |>
  tab_options(
    table.font.size          = 12,
    heading.title.font.size  = 13,
    column_labels.background.color = "#2c3e50",
    column_labels.font.color       = "white"
  )
```

## Figure 6.8 — Numerical Format Comparison {#tbl-numerical}

```{r tbl-numerical}
#| label: tbl-numerical
#| tbl-cap: "The same probability (10%) expressed in six different numerical
#|   formats. Format choice affects risk perception and comprehension.
#|   Adapted from Spiegelhalter (2017)."

tibble(
  Format             = c("Percentage", "Natural frequency",
                         "Decimal probability", "Odds", "1-in-X ratio",
                         "Absolute risk difference"),
  Expression         = c("10%", "10 out of 100", "0.10",
                         "1 to 9 (in favour)", "1 in 10",
                         "+5 per 1,000 (vs baseline of 45 per 1,000)"),
  `Cognitive notes`  = c(
    "Widely understood; may feel abstract for very small probabilities",
    "Intuitive; grounds the probability in a concrete, countable group",
    "Precise; standard in technical reporting; unfamiliar to many lay audiences",
    "Common in gambling contexts; easily misunderstood elsewhere",
    "⚠ Inflates risk perception relative to equivalent N-in-X formats",
    "Most honest for risk changes; always prefer over relative risk"
  ),
  `Best for`         = c(
    "General and informed audiences",
    "General public; medical risk communication",
    "Technical audiences",
    "Specialist contexts only",
    "Avoid — use N-in-X instead",
    "All audiences when communicating risk changes"
  )
) |>
  gt() |>
  tab_header(
    title    = "Numerical Formats for the Same Probability (10%)",
    subtitle = "Adapted from Spiegelhalter (2017) and EFSA (2019)"
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style     = cell_fill(color = "#eaf4fb"),
    locations = cells_body(rows = c(1, 3, 5))
  ) |>
  tab_style(
    style     = cell_fill(color = "#fde8e8"),
    locations = cells_body(rows = 5)
  ) |>
  tab_style(
    style     = cell_text(color = "#c0392b", weight = "bold"),
    locations = cells_body(columns = `Best for`, rows = 5)
  ) |>
  tab_source_note(
    "Key rule: Fix the denominator when comparing frequencies. Avoid 1-in-X formats.
     Always specify the reference class."
  ) |>
  cols_width(
    Format            ~ px(140),
    Expression        ~ px(200),
    `Cognitive notes` ~ px(260),
    `Best for`        ~ px(170)
  ) |>
  tab_options(
    table.font.size          = 11,
    heading.title.font.size  = 13,
    column_labels.background.color = "#2c3e50",
    column_labels.font.color       = "white"
  )
```

# Conclusion Figures {#sec-fig-7}

## Figure 7.1 — Method-Audience Summary Table {#tbl-method-audience}

```{r tbl-method-audience}
#| label: tbl-method-audience
#| tbl-cap: "Summary of the eight uncertainty communication methods reviewed
#|   in this paper, mapped to appropriate audience levels and summarising
#|   each method's primary use case, key strength, and key limitation."

tibble(
  Method = c(
    "Forecast Trajectories",
    "Fan Charts",
    "Distributions",
    "Prediction Intervals & Quantiles",
    "Categorical Approaches",
    "Scenario & What-If Analysis",
    "Verbal Expressions",
    "Numerical Expressions"
  ),
  `Best\naudience level` = c(
    "Technical",
    "Technical",
    "Technical",
    "Entry → Technical",
    "Entry → Informed",
    "Informed → Technical",
    "Entry → Informed",
    "Entry → Technical"
  ),
  `Primary use case` = c(
    "Stochastic variation; cumulative / path-dependent decisions",
    "Time-varying uncertainty with expanding horizon; balance of risks",
    "Full probabilistic detail for expert downstream modelling",
    "Risk-averse decisions; mass-media communication of uncertainty",
    "Meta-uncertainty; unquantifiable confidence in evidence quality",
    "Deep uncertainty; long-horizon strategic planning",
    "Meta-uncertainty; implicit directionality signals",
    "Consequential decisions requiring aggregation of probabilities"
  ),
  `Key strength` = c(
    "Shows true path variability and cumulative uncertainty",
    "Visually encodes widening horizon and distributional skewness",
    "Informationally complete; enables extraction of any quantile",
    "Versatile; preserves trust; directly actionable",
    "Accessible; honest for unquantifiable uncertainty",
    "Avoids false precision under deep uncertainty",
    "Natural; directional; accessible without statistical training",
    "Unambiguous ordering; aggregable; receiver-preferred"
  ),
  `Key limitation` = c(
    "Visual clutter; cherry-picking of single line",
    "Overwhelming for non-experts; deterministic construal error",
    "Cognitively demanding; mode/mean confusion; variability conflation",
    "Uniform distribution fallacy; anchoring on extremes",
    "Information loss; conflates likelihood and confidence",
    "Deterministic misuse; future myopia; simulation heuristic",
    "Massive interpretive variability; severity and extremity effects",
    "Reference class ambiguity; ratio bias; 1-in-X inflation"
  )
) |>
  gt() |>
  tab_header(
    title    = "Uncertainty Communication Methods: A Summary Guide",
    subtitle = "Audience levels, use cases, strengths, and limitations"
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style     = cell_text(weight = "bold", color = "#1a5276"),
    locations = cells_body(columns = Method)
  ) |>
  tab_style(
    style     = cell_fill(color = "#eaf4fb"),
    locations = cells_body(rows = c(1, 3, 5, 7))
  ) |>
  tab_style(
    style     = cell_fill(color = "#fdfefe"),
    locations = cells_body(rows = c(2, 4, 6, 8))
  ) |>
  # Colour-code audience levels
  tab_style(
    style     = cell_text(color = "#c0392b"),
    locations = cells_body(
      columns = `Best\naudience level`,
      rows    = c(1, 2, 3)
    )
  ) |>
  tab_style(
    style     = cell_text(color = "#27ae60"),
    locations = cells_body(
      columns = `Best\naudience level`,
      rows    = c(5, 7)
    )
  ) |>
  cols_width(
    Method                  ~ px(160),
    `Best\naudience level`  ~ px(120),
    `Primary use case`      ~ px(220),
    `Key strength`          ~ px(200),
    `Key limitation`        ~ px(200)
  ) |>
  tab_source_note(
    "Colour coding: Red = Technical audiences only; 
     Green = Suitable for entry-level; Black = Broad applicability."
  ) |>
  tab_options(
    table.font.size          = 11,
    heading.title.font.size  = 13,
    heading.subtitle.font.size = 11,
    column_labels.background.color = "#2c3e50",
    column_labels.font.color       = "white"
  )
```
